---
title: "Peer Assignment 1"
author: "jesussuniaga"
date: "19 de diciembre de 2015"
output: html_document
---
# 1. Loading the data
 

 ```{r}
unzip("activity.zip")
data <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
sinNA <- na.omit(data)
```
# 

#2. What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.

Make a histogram of the total number of steps taken each day

Calculate and report the mean and median total number of steps taken per day
```{r}
attach(sinNA)
totalSteps <- aggregate(steps, list(Date = date), FUN = "sum")
mean(totalSteps$x)
median(totalSteps$x)
hist(totalSteps$x, col = "orangered", xlab = "Number of Steps Taken Each Day", 
     ylab =   "freq", main = "Histogram of Total Number of Steps Taken Each Day")
```

#3.What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
library(ggplot2)
timeseries <- aggregate(steps, list(interval = as.numeric(as.character(interval))), FUN = "mean")
ggplot(timeseries, aes(interval, y = x)) + geom_line() + labs(title = "Time Series Plot of the 5-minute Interval", x = "5-minute intervals", y = "Average Number of Steps Taken")
maximo <- which.max(timeseries$x)
maximo
timeseries [maximo, ]

```
## 4. Imputing missing values


Devise a strategy for filling in all of the missing values in the dataset. 
The strategy does not need to be sophisticated. For example, you could use the mean/median
for that day, or the mean for that 5-minute interval, etc.
My strategy is to use the mean for that 5-minute interval to fill each NA value 
in the steps column.
Now I Create a new dataset that is equal to the original dataset
but with the missing data filled in.
```{r}
newdata <- data 
for (i in 1:nrow(newdata)) {
     if (is.na(newdata$steps[i])) {
        newdata$steps[i] <- timeseries[which(newdata$interval[i] == timeseries$interval), ]$x
                   }
            }

  head(newdata)
sum(is.na(newdata))
```
## 5. Imputing missing values. Second Part.
Make a histogram of the total number of steps taken each day and Calculate 
and report the mean and median total number of steps taken per day. 
Do these values differ from the estimates from the first part of the assignment?
What is the impact of imputing missing data on the estimates of the total daily 
number of steps?
My intrepretation is that The mean value should'not change because I use the mean for that 5-minute interval to fill each NA value in the steps column. And mathematically I can show that the new mean is te same. But other statistics may change a little bit.

```{r}
totalSteps2 <- aggregate(newdata$steps, list(Date = newdata$date), FUN = "sum")
mean(totalSteps2$x)
median(totalSteps2$x)
hist(totalSteps2$x, col = "blue", xlab = "Number of Steps Taken Each Day",
     ylab = "freq",main = "Histogram of Total Number of Steps Taken Each Day" )
```

## 6. Are there differences in activity patterns between weekdays and weekends?
 Creating a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
newdata$date <- as.Date(newdata$date, "%Y-%m-%d")
newdata$day <- weekdays(newdata$date)
happydays <- c("sábado","domingo")

newdata$tipodia<-as.factor(ifelse(weekdays(newdata$date)%in%happydays,"weekend","weekday"))

library(dplyr)
by_tipodia <- group_by(newdata, tipodia)
a <- summarize(by_tipodia, mean(steps))
a
by_tipodia
qplot(interval, steps, data=by_tipodia, geom=c("line"), xlab="5-min intervals", 
      ylab="steps mean", main="") + facet_wrap(~ tipodia, ncol=1)